\texttt{JAXbind}: Bind any function to JAX

### 2024-03-12 paper/paper.tex
paper: tweak
paper.tex: Sync

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={JAXbind: Easy bindings to JAX},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

@@

\title{\texttt{JAXbind}: Easy bindings to JAX}

@@

The use of JAX (Bradbury et al., 2018) is widespread in the natural
sciences.
Of particular interest is JAX's powerful transformation system.
It enables to retrieve arbitrary derivatives of functions, batch computations, and just-in-time code for additional performance.
Its transformation system relies on all components of the computation being written in JAX.

@@

We believe \texttt{JAXbind} to be highly useful in scientific computing.
We intend to use this package to connect the Hartley transform and the
spherical harmonic transform from DUCC (Reinecke, 2024) to the probabilistic programming package NIFTy
(Edenhofer et al., 2024) as well as the radio interferometry response
from DUCC with the radio astronomy package \texttt{resolve} (Arras et al., 2024) for radio
astronomy. Furthermore, we intend to connect the non-uniform FFT from
DUCC with JAX for applications in strong-lensing astrophysics. We
envision many further applications within and outside of astrophysics.

@@

Automatic differentiation is a core feature of JAX and often one of the
main reasons for using it. Thus, it is essential that custom functions
registered with JAX support automatic differentiation. In the following,
we will outline which functions our package respectively JAX requires to
enable automatic differentiation. For simplicity, we assume that we want
to connect the nonlinear function \(f(x_1,x_2) = x_1x_2^2\) to JAX. The
\texttt{JAXbind} package expects the Python function for \(f\) to take
three positional arguments. The first argument, \texttt{out}, is a
\texttt{tuple} into which the results are written. The second
argument is also a \texttt{tuple} containing the input to the function,
in our case, \(x_1\) and \(x_2\). Via \texttt{kwargs\_dump}, any
keyword arguments given to the registered JAX primitive can be
forwarded to $f$ in a serialized form.

@@

To just-in-time compile the function, JAX needs to abstractly evaluate
the code, i.e.~it needs to be able to infer the shape and dtype of the
output of the function given only the shape and dtype of the
input. We have to provide these abstract evaluation functions returning
the output shape and dtype given an input shape and dtype for $f$
as well as for the \texttt{vjp} application. The output shape of the
\texttt{jvp} is identical to the output shape of $f$ itself and
does not need to be specified again.
The abstract evaluation functions take normal positional and keyword arguments.

@@

Currently, \texttt{JAXbind} only supports primitives that act on CPU memory.
In the future, GPU support could be added analogous to the CPU backend.
Note, the automatic differentiation in JAX is backend agnostic and would thus not require any additional bindings.

@@

\hypertarget{references}{%
\section{References}\label{references}}

>>>>

\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={JAXbind: Bind any function to JAX},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}

@@

\title{\texttt{JAXbind}: Bind any function to JAX}

@@

The use of JAX (Bradbury et al., 2018) is widespread in the natural
sciences. Of particular interest is JAX's powerful transformation
system. It enables to retrieve arbitrary derivatives of functions, batch
computations, and just-in-time compile code for additional performance.
Its transformation system relies on all components of the computation
being written in JAX.

@@

We believe \texttt{JAXbind} to be highly useful in scientific computing.
We intend to use this package to connect the Hartley transform and the
spherical harmonic transform from DUCC (Reinecke, 2024) to the
probabilistic programming package NIFTy (Edenhofer et al., 2024) as well
as the radio interferometry response from DUCC with the radio astronomy
package \texttt{resolve} (Arras et al., 2024). Furthermore, we intend to
connect the non-uniform FFT from DUCC with JAX for applications in
strong-lensing astrophysics. We envision many further applications
within and outside of astrophysics.

@@

Automatic differentiation is a core feature of JAX and often one of the
main reasons for using it. Thus, it is essential that custom functions
registered with JAX support automatic differentiation. In the following,
we will outline which functions our package respectively JAX requires to
enable automatic differentiation. For simplicity, we assume that we want
to connect the nonlinear function \(f(x_1,x_2) = x_1x_2^2\) to JAX. The
\texttt{JAXbind} package expects the Python function for \(f\) to take
three positional arguments. The first argument, \texttt{out}, is a
\texttt{tuple} into which the results are written. The second argument
is also a \texttt{tuple} containing the input to the function, in our
case, \(x_1\) and \(x_2\). Via \texttt{kwargs\_dump}, any keyword
arguments given to the registered JAX primitive can be forwarded to
\(f\) in a serialized form.

@@

To just-in-time compile the function, JAX needs to abstractly evaluate
the code, i.e.~it needs to be able to infer the shape and dtype of the
output of the function given only the shape and dtype of the input. We
have to provide these abstract evaluation functions returning the output
shape and dtype given an input shape and dtype for \(f\) as well as for
the \texttt{vjp} application. The output shape of the \texttt{jvp} is
identical to the output shape of \(f\) itself and does not need to be
specified again. The abstract evaluation functions take normal
positional and keyword arguments.

@@

Currently, \texttt{JAXbind} only supports primitives that act on CPU
memory. In the future, GPU support could be added analogous to the CPU
backend. The automatic differentiation in JAX is backend agnostic and
would thus not require any additional bindings to work on the GPU.

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}
